---
title: "Dimensional reduction"
author: "Piotr Szefer"
date: "September 15, 2017"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# What can I do with these framework?

Secondary extinctions? I can generate random community with given distribution of traits. Trophic link strength is something invariant, independent on temporal and spatial dynamics.

# Is the dimensionality reasonable

```{r}
A <- function(a0,tr,tc){
  return(a0*exp(-1/2 * (dist(rbind(tr,tc),method = "euclidean"))^2))
}

# How it is approximated?
a0 <- 0.5
epsilon <- 3/4*a0

# Region where a is never <= a0/4
# Region where is never >= 3/4*a0
# These are harmless, as any value in [0,a0] provides a valid 
# approximiation of these
vals <- seq(0,a0,length=100) # any of this values is a valid approximation of safe
safe <- seq(a0/4, 3/4*a0, length=100)
#vals + epsilon >= safe

# These are the values of circles 
#sqrt(-2*log(1/4)) # values bigger than a0/4
#sqrt(-2*log(3/4)) # values smaller than 3*a0/4

# If i pick any value between 0 and a0 i can be sure that I have approximated
# some value from the region where A is never <= a0/4, and never >= 3*a0/4.
tc <- c(5,5)

rows <- seq(0,10,by=0.1)
cols <- seq(0,10,by=0.1)

M <- matrix(0, nrow = length(rows),
            ncol=length(cols))
for (row in 1:length(rows)){
  for (col in 1:length(cols)){
    M[row,col] <- A(a0, c(rows[row],cols[col]), tc)
  }
}

rownames(M) <- rows
colnames(M) <- cols

# colours <- rep("red",length(M))
# boolM <- (M < (3*a0/4) & M > a0/4)
# as.vector(t(boolM))
# colours[as.vector(boolM)] <- "blue"
# 
# library(rgl)
# persp3d(M, col=colours)



```

Probability distributions or **empirical distributions** of data points over these spaces are assumed given.

Following from Rossberg 2013 i have simulated all the elements necessary to evaluate and reduce the dimentionality of trophic niche spaces. I have build vulnerability $\mathbf{v}^{'}$ and foraging $\mathbf{f}^{'}$ trait vectors sampling from a normal distribution. They were sampled independently. So that $Cov(v_i^{'}, v_j^{'}) = \delta_{j,k}$.Vector $\mathbf{x}$ is a result of concatenation of vulnerability and foraging trait vectors:
$$\mathbf{x} = \begin{pmatrix} \mathbf{v}^{'} \\ \mathbf{f}^{'} \end{pmatrix}$$
As a link function following equation will be used:
$$ A^{'}(\mathbf{v}^{'}, \mathbf{f}^{'}) = a_0^{'} \exp{ \left[ \mathbf{b}^{T}\mathbf{x} + \frac{1}{2}\mathbf{x}^T\mathbf{C}\mathbf{x} \right] } $$
Symmetric matrix $\mathbf{C}$ and vector of constants $\mathbf{b}$ would have to be calculated from the empirical data of traits and link strengths.

```{r}
# Food webs and diversity

Dprime <- 3
v <- rnorm(Dprime) # Species R
f <- rnorm(Dprime) # Species C

x <- c(v,f)

# I don tthink this is the community matrix, it is just the parameters matrix
# which needs to be estimated.
# To construct C matrices consult deRuiter and Fath 2007

C_vals <- runif(((Dprime*2)^2 - Dprime*2)/2, min=-1,max=1)
C <- matrix(0, nrow=Dprime*2, ncol=Dprime*2)
C[upper.tri(C)] <- C_vals
C <- C + t(C)
diag(C) <- 1

b <- c(rep(1,Dprime+Dprime))

# Link strength between RC:
A <- function(a0 = 1, b, x, C){
  link_str <- a0 * exp(t(b) %*% x + 1/2 * t(x) %*% C %*% x)
  return(link_str)
}
```

```{r}
# Decomposition of the equation:
EC <- eigen(C)
lambda <- EC$values
u <- EC$vectors

# Spectral value decomposition
Crec <- matrix(0, nrow=Dprime*2, ncol=Dprime*2)
Crec <- u %*% diag(lambda) %*% t(u)
```

Matrix C was decomposed using spectral value decomposition. And concatenated trait vectors $x$ were represented using eigenvectors of matrix C.

$$ A^{'}(\mathbf{v}^{'}, \mathbf{f}^{'}) = a_0^{'} \exp{ \left[ \sum_{k=1}^{D^{'}} (\mathbf{b}^T\mathbf{u}_k)^{T}\mathbf{u}^T_k\mathbf{x} + \frac{1}{2}\sum_{k=}^{D^{'}}\lambda_k(\mathbf{u}^T_k\mathbf{x})^2 \right] }  $$
Lets see what is going to happen if we use only few traits

```{r}

Adim <- function(a0 = 1, b,x,C, traits=10){
  EC <- eigen(C)
  EC <- eigen(C)
  lambda <- EC$values
  return(a0*exp((t(b) %*% u) %*% (t(u) %*% x) + 1/2 * (t(x)%*%u%*%diag(lambda)%*%t(u)%*%x)))
}

```

Covariance matrix: $\sum = \mathbb{E} \left[ (\mathbf{X} - \mathbb{E} [\mathbf{X}] )(\mathbf{X} - \mathbb{E} [\mathbf{X}] )^{'} \right]$

```{r}
v = rnorm(10, 5,1)
f = rnorm(10, 1,2)
# Covariate matrix
covM = function(v){1/length(v)*(v - mean(v)) %*% t(v - mean(v))}
Q = covM(v - f)
cov(v - f %*% t(v - f))
```
